Equation-of-state insensitive relations after GW170817

### 2019-02-24 paper/bin-love-after-gw170817.tex
katerina comments conclusions

\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{stackedFisher.eps}
\end{center}
\caption{(Color online) Fisher-estimated statistical uncertainties on the extraction of $\lambda_{0}$ with interferometer (O2, aLIGO, A\texttt{+}, Voyager, CE, ET-D) as a function of the signal-to-noise-ratio expected in each of these instruments, given a single GW170817 detection (circles). The statistical uncertainties with ET are lower than with CE in spite of a lower signal-to-noise ratio because the former is more sensitive above 300 Hz, where tidal effects matter the most (see Sec.~\ref{sec:futureObservations} for further discussion). We also plot the combined statistical uncertainty given $N$ observations consistent with the NS binary merger rate for a 1 year observation (regions), with the top and bottom edges of the regions corresponding to optimistic and pessimistic merger rates. These statistical uncertainties should be compared to the systematic uncertainty on the extraction of $\lambda_{0}$ due to EoS-variability (horizontal dashed line). Observe that the statistical and systematic uncertainties cross for Voyager-class detectors. We confirm this conclusion by repeating the statistical analysis with two different waveform models (PhenomD~\cite{PhenomDI,PhenomDII} plus NRTidal corrections~\cite{Samajdar:NRTidal} and PhenomD~\cite{PhenomDI,PhenomDII} plus 6PN tidal corrections~\cite{Wade:tidalCorrections}).}
\label{fig:stackedFisher}
\end{figure}

@@

The left panel of Fig.~\ref{fig:clove} shows the C-Love relations for both the constrained and unconstrained sets, along with the corresponding \emph{absolute differences} from the fits (instead of the fractional differences as done back in Fig.~\ref{fig:ILQ}). Observe how the fit to the constrained EoSs suppresses the EoS variability compared to the fit to the unconstrained set, as well as that of previous works~\cite{Yagi:binLove}. The maximal EoS variation is compared between these three cases in Table~\ref{tab:maxVar}.

@@

Figures~\ref{fig:singleFisherLt} and~\ref{fig:stackedFisher}, and Table~\ref{tab:variances} summarize our results graphically using an IMRD + 6PN template. In the latter, you can see the statistical accuracy to which $\tilde{\Lambda}$ can be estimated as a function of $\rho^A_{\text{GW170817}}$ for single GW170817-like events, i.e.~the signal-to-noise ratio that future detector $A$ would have measured for a GW170817 event. Observe that the systematic uncertainty in $\tilde{\Lambda}$ (the horizontal dashed line) becomes comparable to the statistical uncertainty for Voyager-class detectors or better. Figure~\ref{fig:singleFisherLt} presents similar results but for the parameter $\lambda_{0}$, for which one can combine posteriors and obtained a reduced combined statistical error. This figure also shows that the statistical and systematic uncertainties become comparable for Voyager-class detectors or better. The careful reader will notice that Figs.~\ref{fig:stackedFisher} and~\ref{fig:singleFisherLt} show that the single-event statistical uncertainties using CE is higher than using ET despite the later having a larger SNR for a GW170817-like event. This is because the 3-detector geometry of ET enables it to have higher sensitivity than ET at frequencies above 300Hz, which is precisely where the tidal deformabilities is encoded, but ET has lower sensitivity at lower frequencies, where a lot of the signal-to-noise ratio accumulates.

\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{singleFisherLt.eps}
\end{center}
\caption{(Color Online) Estimated statistical uncertainty $\sigma^A_{\text{GW170817}}$ in the extraction of $\tilde\Lambda$ from a single GW170817-like event as if observed with aLIGO at design sensitivity, A\texttt{+}, Voyager, CE, and ET-D, plotted as a function of the signal-to-noise-ratio $\rho^A_{\text{GW170817}}$ that those detectors would measure for such an event. For comparison, we also plot the systematic uncertainty on $\tilde\Lambda$ due to the use of the binary Love relations. Observe that the statistical and the systematic errors become comparable for Voyager-class detectors or better.
}
\label{fig:singleFisherLt}
\end{figure}

@@

But is the statistical error calculated here robust to mismodeling systematics in the template? We repeat the analysis above using an IMRD + NRTidal template model  and find results consistent with those presented above. Indeed, Fig.~\ref{fig:stackedFisher} shows the individual and combined statistical error on $\lambda_{0}$ when using this template model (maroon circles and dashed region). Observe that the accuracy to which $\lambda_{0}$ can be measured with this template model is systematically better than when using the IMRD+6PN model (by roughly a factor of two). This implies that more accurate template models will indeed be required in the third-generation detector era, as previously pointed out in~\cite{Samajdar:NRTidal}. However, our conclusion that for Voyager-class detectors or better the statistical and binary Love systematic uncertainties become comparable seems robust.

>>>>

\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{stackedFisher.eps}
\end{center}
\caption{(Color online) Fisher-estimated statistical uncertainties on the extraction of $\lambda_{0}$ with interferometer (O2, aLIGO, A\texttt{+}, Voyager, CE, ET-D) as a function of the signal-to-noise-ratio expected in each of these instruments, given a single GW170817 detection (circles). The statistical uncertainties with ET are lower than with CE in spite of a lower signal-to-noise ratio because the former is more sensitive above 300 Hz, where tidal effects matter the most (see Sec.~\ref{sec:futureObservations} for further discussion). We also plot the combined statistical uncertainty given $N$ observations consistent with the NS binary merger rate for a 1 year observation (regions), with the top and bottom edges of the regions corresponding to optimistic and pessimistic merger rates. These statistical uncertainties should be compared to the systematic uncertainty on the extraction of $\lambda_{0}$ due to EoS-variability (horizontal dashed line). The statistical and systematic uncertainties cross for Voyager-class detectors. We confirm this conclusion by repeating the statistical analysis with two different waveform models (PhenomD~\cite{PhenomDI,PhenomDII} plus NRTidal corrections~\cite{Samajdar:NRTidal} and PhenomD~\cite{PhenomDI,PhenomDII} plus 6PN tidal corrections~\cite{Wade:tidalCorrections}).}
\label{fig:stackedFisher}
\end{figure}

@@

The left panel of Fig.~\ref{fig:clove} shows the C-Love relations for both the constrained and unconstrained sets, along with the corresponding \emph{absolute differences} from the fits (instead of the fractional differences as done back in Fig.~\ref{fig:ILQ}). The fit to the constrained EoSs suppresses the EoS variability compared to the fit to the unconstrained set, as well as that of previous works~\cite{Yagi:binLove}. The maximal EoS variation is compared between these three cases in Table~\ref{tab:maxVar}.

@@

Figures~\ref{fig:singleFisherLt} and~\ref{fig:stackedFisher}, and Table~\ref{tab:variances} summarize our results graphically using an IMRD + 6PN template. In the latter, you can see the statistical accuracy to which $\tilde{\Lambda}$ can be estimated as a function of $\rho^A_{\text{GW170817}}$ for single GW170817-like events, i.e.~the signal-to-noise ratio that future detector $A$ would have measured for a GW170817 event. The systematic uncertainty in $\tilde{\Lambda}$ (the horizontal dashed line) becomes comparable to the statistical uncertainty for Voyager-class detectors or better. Figure~\ref{fig:singleFisherLt} presents similar results but for the parameter $\lambda_{0}$, for which one can combine posteriors and obtained a reduced combined statistical error. This figure also shows that the statistical and systematic uncertainties become comparable for Voyager-class detectors or better. The careful reader will notice that Figs.~\ref{fig:stackedFisher} and~\ref{fig:singleFisherLt} show that the single-event statistical uncertainties using CE is higher than using ET despite the later having a larger SNR for a GW170817-like event. This is because the 3-detector geometry of ET enables it to have higher sensitivity than ET at frequencies above 300Hz, which is precisely where the tidal deformabilities is encoded, but ET has lower sensitivity at lower frequencies, where a lot of the signal-to-noise ratio accumulates.

\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{singleFisherLt.eps}
\end{center}
\caption{(Color Online) Estimated statistical uncertainty $\sigma^A_{\text{GW170817}}$ in the extraction of $\tilde\Lambda$ from a single GW170817-like event as if observed with aLIGO at design sensitivity, A\texttt{+}, Voyager, CE, and ET-D, plotted as a function of the signal-to-noise-ratio $\rho^A_{\text{GW170817}}$ that those detectors would measure for such an event. For comparison, we also plot the systematic uncertainty on $\tilde\Lambda$ due to the use of the binary Love relations. The statistical and the systematic errors become comparable for Voyager-class detectors or better.
}
\label{fig:singleFisherLt}
\end{figure}

@@

But is the statistical error calculated here robust to mismodeling systematics in the template? We repeat the analysis above using an IMRD + NRTidal template model  and find results consistent with those presented above. Indeed, Fig.~\ref{fig:stackedFisher} shows the individual and combined statistical error on $\lambda_{0}$ when using this template model (maroon circles and dashed region). The accuracy to which $\lambda_{0}$ can be measured with this template model is systematically better than when using the IMRD+6PN model (by roughly a factor of two). This implies that more accurate template models will indeed be required in the third-generation detector era, as previously pointed out in~\cite{Samajdar:NRTidal}. However, our conclusion that for Voyager-class detectors or better the statistical and binary Love systematic uncertainties become comparable seems robust.
